### 2021-10-19 text/outline/outline.tex
update notation of thresholding theory section

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we set the perturbation probability $\pi$ to $1/2$. Additionally, without loss of generality, we set the target of inference $\beta^m_1$ to $1$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we state several results relating the limiting bias $b$ of $\hat{\beta}^m_1$ to $c_\textrm{bayes}$ and $\beta^g_1$. We defer proofs to the appendix.

@@

This section contains proofs for results presented the section ``Theoretical analysis of thresholding estimator.''

\subsubsection{Limit of $\hat{\beta}^m_1$ in Gaussian model with intercepts}

Consider the thresholding estimator $\hat{\beta}^m_1$ for the Gaussian model with intercepts (\ref{theoretical_model}). Dividing by $n$ in (\ref{thresh_est_intercepts}), we can express $\hat{\beta}^m_1$ as
$$ \hat{\beta}^m_1 = \frac{ \frac{1}{n} \sum_{i=1}^n ( \hat{p}_i - \overline{\hat{p}_i})(m_i - \overline{m})}{ \frac{1}{n} \sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$ By WLLN,
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\textrm{Cov}(\hat{p}_i, m_i)}{\V\left(\hat{p}_i\right)}.$$ To compute this quantity, we first compute several simpler quantities:
\begin{itemize}
\item[1.] \textbf{Expectation of $m_i$}: $\E[m_i] = \beta^m_0 + \beta^m_1\pi.$
\item[2.] \textbf{Expectation of $\hat{p}_i$}: Let $f$ denote the $N(0,1)$ density. Then \begin{multline*}
\E[\hat{p}_i] = \P\left[\hat{p}_i = 1\right] = \P\left[\beta^g_0 + \beta^g_1 + \tau_i \geq c \right] = \\ \textrm{(By LOTP) } \P\left[ \beta^g_0 + \tau_i \geq c \right]\P\left[p_i = 0\right] + \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \P[p_i = 1] \\ = \P\left[ \tau_i \geq c - \beta^g_0\right](1- \pi) + \P\left[ \tau_i \geq c - \beta^g_1 - \beta^g_0 \right](\pi) \\ = \left(\int_{c - \beta^g_0}^\infty f\right) (1 - \pi) + \left( \int_{c - \beta^g_1 - \beta^g_0}^\infty f \right)(\pi) = \zeta(1-\pi) + \omega \pi,
\end{multline*}
where $$\zeta = \int_{c - \beta_0^g}^\infty f = 1 - \Phi(c - \beta^g_0)$$ and 
$$ \omega = \int_{c - \beta^g_1 - \beta^g_0} f = 1 - \Phi(c - \beta^g_1 - \beta^g_0).$$
\item[3.] \textbf{Expectation of $\hat{p}_i p_i$}: 
$$\E\left[ \hat{p}_i p_i \right] = \E\left[ \hat{p}_i | p_i = 1 \right] \P\left[ p_i =1 \right] = \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \pi = \omega \pi.$$
\item[4.] \textbf{Expectation of $\hat{p}_i m_i$}:
\begin{multline*}
\E\left[\hat{p}_i m_i\right] = \E[\hat{p}_i (\beta^m_0 + \beta^m_1 p_i + \ep_i)] = \beta^m_0 \E\left[\hat{p}_i\right] + \beta^m_1 \E\left[\hat{p}_i p_i\right] + \E[\hat{p}_i \ep_i] \\ = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi + \E[\hat{p}_i] \E[\ep_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi.
\end{multline*}
\item[5.] \textbf{Variance of $\hat{p}_i$}: Because $\hat{p}_i$ is binary, we have that $\V[\hat{p}_i] = \E[\hat{p}_i]\left(1 - \E[\hat{p}_i]\right) .$
\item[6.] \textbf{Covariance of $\hat{p}_i, m_i$}:
\begin{multline*}
\textrm{Cov}\left(\hat{p}_i, m_i\right) = \E\left[\hat{p}_i m_i\right] - \E[\hat{p}_i] \E[m_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi - \E[\hat{p}_i]( \beta^m_0 + \beta^m_1 \pi)\\ = \beta^m_1 \omega \pi - \E[\hat{p}_i] \beta_1^m \pi = \beta^m_1 \pi \left( \omega - \E[\hat{p}_i]\right).
\end{multline*}
Combining these expressions, we have that
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\beta^m_1 \pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i](1 - \E[\hat{p}_i])} = \beta^m_1 \gamma(\beta^g_0, \beta^g_1, c, \pi).$$
\end{itemize}
\subsubsection{Re-expressing $\gamma$ when $\pi = 1/2$}

Suppose $\pi = 1/2$. We rewrite the attenuation fraction $\gamma$ in a way that makes it more amenable to theoretical analysis.

>>>>

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we begin by setting the perturbation probability $\pi$ to $1/2$. Additionally, without loss of generality, we set the target of inference $\beta^m_1$ to $1$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we state several results relating the limiting bias $b$ of $\hat{\beta}^m_1$ to $c_\textrm{bayes}$ and $\beta^g_1$. We defer proofs to the appendix.

@@

This section contains proofs for results presented the section ``Theoretical analysis of thresholding estimator.'' The subsections are arranged as follows: FILL IN...

\subsubsection{Notation}
 All notation introduced here pertains to the Gaussian model with intercepts (\ref{theoretical_model}). Recall that the attenuation function $\gamma: \R^4 \to \R$ is defined by
$$ \gamma(\beta^g_1, c, \pi, \beta^g_0) = \frac{\pi(\omega - \E[\hat{p}_i])}{ \E[\hat{p}_i](1 -\E[\hat{p}_i])},$$ where $$\begin{cases}
\omega = 1 - \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = 1 - \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$ Additionally, recall that the bias function $b: \R^5 \to \R$ is defined by
$$ b(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1) = \beta^m_1 (1 - \gamma(\beta^g_1, c, \pi, \beta^g_0)).$$ The parameters $\beta^g_0$ and $\beta^m_1$ will be given, fixed constants throughout the proofs. Therefore, to minimize notation, we typically use $\gamma(\beta^g_1, c, \pi)$ (resp., $b(\beta^g_1, c, \pi)$) to refer to the function $\gamma$ (resp., $b$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$ (resp., $(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1)$). Next, we define the functions $g$ and $h: \R^4 \to \R$ by
\begin{equation}\label{def_g}
g(\beta^g_1, c, \pi, \beta^g_0) = (1-\pi)\left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right)\end{equation}
 and
\begin{multline}\label{def_h}
h(\beta^g_1, c, \pi, \beta^g_0) = \left[(1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi\left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \cdot \\ \left[(1-\pi)\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi\left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right],
\end{multline} where $f: \R \to [0,\infty)$ is the $N(0,1)$ density. We will refer to $g$ and $h$ repeatedly throughout the proofs. Similarly to above, we typically use the notation $g(\beta^g_1, c, \pi)$ (resp.\ $h(\beta^g_1, c, \pi)$) to refer to $g$ (resp.\ $h$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$.

Finally, for a given function $r: \R^{p} \to \R$, point $x \in \R^p$, and index $i \in \{1, \dots, p\}$, we use the symbol $D_i r(x)$ to refer to the derivative of the $i$th component of $r$ evaluated at $x$ (\textit{sensu} \cite{fitzpatrick2009}). For example, $D_1 \gamma(\beta^g_1, c, 1/2)$ is the derivative of the first component of $\gamma$ (the component corresponding to $\beta^g_1$) evaluated at $(\beta^g_1, c, 1/2)$. Likewise,  $D_2g(\beta^g_1, c, \pi)$ is the derivative of the second component of $g$ (the component corresponding to $c$) evaluated at $(\beta^g_1, c, \pi).$

\subsubsection{Limit of $\hat{\beta}^m_1$}

We derive the limit in probability of $\hat{\beta}^m_1$ for the Gaussian model with intercepts (\ref{theoretical_model}). Dividing by $n$ in (\ref{thresh_est_intercepts}), we can express $\hat{\beta}^m_1$ as
$$ \hat{\beta}^m_1 = \frac{ \frac{1}{n} \sum_{i=1}^n ( \hat{p}_i - \overline{\hat{p}_i})(m_i - \overline{m})}{ \frac{1}{n} \sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$ By weak LLN,
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\textrm{Cov}(\hat{p}_i, m_i)}{\V\left(\hat{p}_i\right)}.$$ To compute this quantity, we first compute several simpler quantities:
\begin{itemize}
\item[1.] Expectation of $m_i$: $\E[m_i] = \beta^m_0 + \beta^m_1\pi.$
\item[2.] Expectation of $\hat{p}_i$: Let $f$ denote the $N(0,1)$ density. Then \begin{multline*}
\E[\hat{p}_i] = \P\left[\hat{p}_i = 1\right] = \P\left[\beta^g_0 + \beta^g_1 + \tau_i \geq c \right] = \\ \textrm{(By LOTP) } \P\left[ \beta^g_0 + \tau_i \geq c \right]\P\left[p_i = 0\right] + \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \P[p_i = 1] \\ = \P\left[ \tau_i \geq c - \beta^g_0\right](1- \pi) + \P\left[ \tau_i \geq c - \beta^g_1 - \beta^g_0 \right](\pi) \\ = \left(\int_{c - \beta^g_0}^\infty f\right) (1 - \pi) + \left( \int_{c - \beta^g_1 - \beta^g_0}^\infty f \right)(\pi) = \zeta(1-\pi) + \omega \pi.
\end{multline*}
\item[3.] Expectation of $\hat{p}_i p_i$: 
$$\E\left[ \hat{p}_i p_i \right] = \E\left[ \hat{p}_i | p_i = 1 \right] \P\left[ p_i =1 \right] = \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \pi = \omega \pi.$$
\item[4.] Expectation of $\hat{p}_i m_i$:
\begin{multline*}
\E\left[\hat{p}_i m_i\right] = \E[\hat{p}_i (\beta^m_0 + \beta^m_1 p_i + \ep_i)] = \beta^m_0 \E\left[\hat{p}_i\right] + \beta^m_1 \E\left[\hat{p}_i p_i\right] + \E[\hat{p}_i \ep_i] \\ = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi + \E[\hat{p}_i] \E[\ep_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi.
\end{multline*}
\item[5.] Variance of $\hat{p}_i$: Because $\hat{p}_i$ is binary, we have that $\V[\hat{p}_i] = \E[\hat{p}_i]\left(1 - \E[\hat{p}_i]\right) .$
\item[6.] Covariance of $\hat{p}_i, m_i$:
\begin{multline*}
\textrm{Cov}\left(\hat{p}_i, m_i\right) = \E\left[\hat{p}_i m_i\right] - \E[\hat{p}_i] \E[m_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi - \E[\hat{p}_i]( \beta^m_0 + \beta^m_1 \pi)\\ = \beta^m_1 \omega \pi - \E[\hat{p}_i] \beta_1^m \pi = \beta^m_1 \pi \left( \omega - \E[\hat{p}_i]\right).
\end{multline*}
Combining these expressions, we have that
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\beta^m_1 \pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i](1 - \E[\hat{p}_i])} = \beta^m_1 \gamma(\beta^g_1, c, \pi).$$
\end{itemize}

\subsubsection{Re-expressing $\gamma$ in a simpler form}

We rewrite the attenuation fraction $\gamma$ in a way that makes it more amenable to theoretical analysis. We leverage the fact that $f$ integrates to unity and is even.
First,
\begin{multline}\label{thm:gamma_expression_1} \E\left[\hat{p}_i\right] = (1 - \pi) \left( \int_{c - \beta^g_0}^\infty f \right) + \pi \left(\int_{c - \beta_0^g - \beta_1^g}^\infty f \right) \\ = (1 - \pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right), \end{multline}
 and so \begin{multline}\label{thm:gamma_expression_2} 1 - \E\left[\hat{p}_i\right] = (1 - \pi) + \pi - \E[\hat{p}_i]  = (1-\pi) \left(1 - \int_{c - \beta_0^g}^\infty f \right)  + \pi \left(1 - \int_{c - \beta^g_0 - \beta^g_1}^\infty f \right) \\ = (1 - \pi) \left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi \left( \int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right).
\end{multline}
Next,
\begin{equation}\label{thm:gamma_expression_3}
\omega = \int_{c - \beta^g_1 - \beta^g_0}^\infty f,\end{equation} and so
\begin{multline}\label{thm:gamma_expression_4}
\omega - \E[\hat{p}_i]  = \left( \int_{c - \beta^g_1 - \beta^g_0}^\infty f \right) - (1 - \pi) \left( \int_{c - \beta^g_0}^\infty f \right) - \pi \left( \int_{c - \beta_0^g - \beta_1^g}^\infty f \right) \\ = (1 - \pi) \left(\int_{c - \beta^g_0 - \beta^g_1 }^\infty f \right) - (1- \pi) \left(\int_{c - \beta^g_0}^\infty f \right) \\ =  (1 - \pi)\left(\int_{-\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1 - \pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right).
\end{multline}
Combining (\ref{thm:gamma_expression_1}, \ref{thm:gamma_expression_2}, \ref{thm:gamma_expression_3}, \ref{thm:gamma_expression_4}), we find that
\begin{multline}\label{gamma_alternative}
\gamma(\beta^g_1, c, \pi) = \frac{\pi(\omega - \E[\hat{p}_i])}{\E[\hat{p}_i](1 - \E[\hat{p}_i])} \\ = \frac{\pi \left[ (1 - \pi) \left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1 - \pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right) \right]}{\left[(1-\pi) \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \left[(1 - \pi) \left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi \left( \int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right]}.
\end{multline}
As a corollary, when $\pi = 1/2$, we have that
\begin{multline}\label{gamma_alternative_pi_half}
\gamma(\beta^g_1, c, 1/2)  \\ = \frac{\left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - \left( \int_{-\infty}^{ \beta_0^g - c} f \right) }{ \left[ \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \left[\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right]}.
\end{multline}
Recalling the definitions of $g$ (\ref{def_g}) and $h$ (\ref{def_h}), we can write $\gamma$ as
$$ \gamma(\beta^g_1, c, \pi) = \frac{g(\beta^g_1, c, \pi)}{h(\beta^g_1, c,\pi)}.$$

\subsubsection{Derivatives of  $g$ and $h$ in $c$}

We compute the derivatives of $g$ and $h$ in $c$, which we will need for subsequent results.

% \subsubsection{Bayes-optimal decision boundary as a critical value}
% Let $c_\textrm{bayes} = \beta^g_0 + (1/2)\beta^g_1.$ We show that $c = c_\textrm{bayes}$ is a critical value of $\gamma$ for $\pi = 1/2$ and fixed $(\beta^g_0, \beta^g_1)$ , i.e, $$\frac{\partial  \gamma(\beta^g_0, \beta^g_1, c, 1/2)}{ \partial c} \vline_{[c = c_\textrm{bayes}]} = 0.$$ By quotient rule,
% $$ \frac{\partial  \gamma(\beta^g_0, \beta^g_1, c, 1/2)}{ \partial c} = \frac{g'(\cdot, c) h(\cdot, c) - g(\cdot, c) h'(\cdot, c)}{\left[ h(\cdot, c) \right]^2}.$$ First,
% $$g'(\cdot, c) = - f( \beta^g_0 + \beta^g_1 - c ) + f(\beta^g_0 - c).$$ Second, 
% \begin{multline*}
% h'(\cdot, c) = -[f(\beta^g_0 - c)  + f( \beta^g_0 + \beta^g_1 - c )]  \left[\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right] \\ +[ f(c - \beta^g_0) + f(c - \beta^g_0 - \beta^g_1) ] \left[ \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right].
% \end{multline*}
% Evaluating $g'(\cdot, c)$ and $h'(\cdot, c)$ at $c = c_\textrm{bayes} $, we find that
% $$g'(\beta^g_0 + (1/2) \beta^g_1) = -f(\beta^g_1/2) + f(- \beta^g_1/2) = -f(\beta^g_1/2) + f(\beta^g_1/2) = 0,$$ where the second equality follows from the evenness of $f$. Similarly, 
% \begin{multline*}
% h'(\beta^g_0 + (1/2) \beta^g_1) = - \left[ f( -\beta^g_1/2) + f( \beta^g_1/2 ) \right] \left[ \int_{-\infty}^{ \beta^g_1/2} f + \int_{-\infty}^{ -\beta^g_1/2} f \right] + \\ \left[ f( \beta^g_1/2) + f( -\beta^g_1/2 ) \right] \left[ \int_{-\infty}^{ -\beta^g_1/2} f + \int_{-\infty}^{ \beta^g_1/2 f} \right] = 0.
% \end{multline*}
% We conclude that
% $$ \frac{\partial \gamma(\beta^g_0, \beta^g_1, c, 1/2)}{ \partial c} \vline_{[c = c_\textrm{bayes}]} = \frac{0\cdot h(\cdot, c_\textrm{bayes}) - g(\cdot, c_\textrm{bayes}) \cdot 0}{ h(\cdot, c_\textrm{bayes})^2} = 0.$$ Finally, because asymptotic bias $b$ can be expressed in terms of $\gamma$ as $$b(\beta^g_0. \beta^g_1, c, \pi) = \beta^m_1 \left[ 1 - \gamma(\beta^g_0, \beta^g_1, c, \pi) \right],$$ it follows that $c = c_\textrm{bayes}$ is a critical value of $b$ for $\pi = 1/2$ and fixed $(\beta^g_0, \beta^g_1)$.

% \subsubsection{Limit in $\beta^g_1$}

% We compute the limit of $\gamma(\beta^g_0, \beta^g_1, c, 1/2)$ in $\beta^g_1.$ We have that

% \begin{multline*}
 %\lim_{\beta^g_1 \to \infty } \gamma(\beta^g_0, \beta^g_1, c, 1/2) = \frac{ 1 - \int_{-\infty}^{\beta^g_0 - c } f}{\left[ 1 + \int_{-\infty}^{ \beta^g_0 - c} f \right] \left[\int_{-\infty}^{ c - \beta^g_0} f \right]} \\ = \frac{\int_{-\infty}^{c - \beta^g_0} f }{ \left[1 + \int_{-\infty}^{ \beta^g_0 - c} f \right] \left[\int_{-\infty}^{ c - \beta^g_0} f \right]} = \frac{1}{1 + \int_{-\infty}^{\beta^g_0 - c} f} < 1.
% \end{multline*}

% Similarly,
% \begin{multline*}
% \lim_{\beta^g_1 \to -\infty} \gamma(\beta^g_0, \beta^g_1, c, 1/2) = \frac{ -\int_{-\infty}^{\beta^g_0 - c} f }{\left[ \int_{-\infty}^{\beta^g_0 - c} f \right] \left[ \int_{-\infty}^{c - \beta^g_0} f + 1 \right]} = \frac{-1}{1 + \int_{-\infty}^{c - \beta^g_0} f} > -1.
% \end{multline*}

% \subsubsection{Monotonicity in $\beta^g_1$}
% We prove that $\gamma$ is monotonic in $\beta_1^g$ for given $(\beta^g_0, c)$ and $\pi = 1/2$. Recall that
% $$\gamma(\beta^g_0, \beta^g_1, c, 1/2) = \frac{ g(\beta^g_1, c) }{ h(\beta^g_1, c)}.$$ By the quotient rule,
% $$D_2 \gamma (\beta^g_0, \beta^g_1, c, 1/2) = \frac{D_1 g( \beta^g_1, c) h(\beta^g_1, c) - g(\beta^1_g, c) D_1 h(\beta^g_1, c) }{ [ h(\beta^g_1, c)^2]}.$$ The derivative $D_1 g(\beta^g_1, c)$ of $g$ is 
% $$D_1 g(\beta^g_1, c) = f( \beta^g_0 + \beta^g_1 - c ).$$
